---
title: "Toronto Property Crime Analysis: Changes in Rates and Types of Property Crimes (2014-2023)" 
author: 
  - Jimin Lee
thanks: "Code and data are available at: https://github.com/jamiejiminlee/Toronto-Reported-Crimes.git"
date: 27 September 2024
date-format: long
abstract: "This paper aims to explore Toronto's property crime rates and types of property crimes from 2014 to 2023. The analysis reveals a spike in property crimes in 2023, with shifts towards crime types such as break-and-enter and auto thefts."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(opendatatoronto)
library(knitr)
library(dplyr)
library(readr)
```


# Introduction

In 2023, Toronto's property crime rate rose by 22.28% from 2022, reaching a total of 189,977 recorded incidents, breaking the highest recorded value of property crimes in the city.@Statista2023 This paper explores and compares the rates of property crime across different
geographic divisions of the City of Toronto. 


# Data {#sec-data}

The dataset used for this paper is "Police Annual Statistical Report" (@propertycrime_data), accessed through the Open Data Toronto portal (@opendatatoronto). The dataset was simulated, downloaded, cleaned, and analyzed using the R programming language (@citeR). Despite the existence of other datasets in OpenDataToronto such as "Police Annual Statistical Report - Victims of Crimes", it wasn't chosen for this analysis as the paper focuses solely on property crimes. 

```{r}
#| label: codechunk1
#| fig-cap: "Figure1: Graph compares the number of property crimes from 2013 to 2024, in 1-year increments. Figure displays a sudden spike in property crimes in 2023 - we will be analyzing the data from 2023 to compare different types of property crimes that occured"
#| echo: false

cleaned_data |>
  dplyr::filter(report_year >= 2013 & report_year <= 2024) |>
  ggplot(mapping = aes(x = report_year, y = count)) +
  geom_bar(stat = "identity", fill = "purple") +
  scale_x_continuous(breaks = seq(2013, 2024, by = 1)) +
  labs(
    x = "Year",
    y = "Reported Property Crimes"
  ) 

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it. 



# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.


### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```




# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```



\newpage


# References


